---
title: "Government Expenditure on Education"
author: "Raymond Hung"
format: 
  html: 
    toc: true
    fig-width: 12
    page-layout: full
    self-contained: true
    fig-height: 8
    number-sections: true
    highlight-style: github
    
editor_options: 
  chunk_output_type: console
---

# Library

```{r Lib, warning=FALSE, echo=TRUE, output=FALSE, error=FALSE}
#| code-fold: true
library(tidyverse)
library(rio)
library(here)
library(sjmisc)
library(sjlabelled)
library(janitor)
library(broom)
library(magrittr)
library(gt)
```


# Data

```{r echo=TRUE, message=FALSE, warning=FALSE, output=FALSE}
edu.cost <- read_csv(here("Project 1/01. Data/education expense total.csv"), 
                     skip = 0) |> as_tibble()
edu.cost.meta <- read_csv(here("Project 1/01. Data/education expense total_Metadata_Country.csv"), 
                          skip = 0) |> as_tibble()
```

```{r}
edu.cost      |> dim() # 266 rows
edu.cost.meta |> dim() # 265 rows

code_1 <- edu.cost |> select(country_code)
code_2 <- edu.cost.meta |> select(country_code)

code_1 |> bind_rows(code_2) |> 
  group_by(country_code) |> count() |> filter(n == 1)
# Country code "INX" only shows in edu.cost
# After checking the data, the whole row of INX is empty in edu.cost so it is manually deleted.

edu.cost <- edu.cost |> filter(country_code != "INX")
edu.cost |> dim() # 265 rows

edu.cost.long <- edu.cost |> 
    pivot_longer(cols = `1960` : `2021`,
                 names_to = "Year",
                 values_to = "Total_Expense")
```

# Cleaning and Combining

```{r}
cost <- edu.cost.long |> left_join(y = edu.cost.meta, by = "country_code")
cost <- cost |> relocate(c(region, income_group), .after = country_code)

# Indicator name, Indicator code, table name are redundant so they are removed. 
cost <- cost |> select(-indicator_name, -indicator_code, -table_name)

cost <- cost |> 
    mutate(income_group = factor(income_group,
                                 labels = c("Low income", "Lower middle income", 
                                            "Upper middle income", "High income"),
                                 levels = c("Low income", "Lower middle income", 
                                            "Upper middle income", "High income")))
```

# Browsing

## Missing Value
```{r}
naniar::miss_var_summary(cost)
# There are missing value in the variable "region" and "income group" which needs to check.
```

### Region

```{r}
cost |> filter(is.na(region)) |> distinct(country_name)
# 48 country_name have NA in regions. 
# After examination, they are a region instead of an independent country.
# One of them is the world data
```

### Income Group

```{r}
cost |> filter(is.na(income_group)) |> 
  group_by(is.na(region)) |> 
  distinct(country_name) |> 
  count()
# 49 countries do not have region information, and one of them is not a region.

cost |> filter(is.na(income_group) & !is.na(region)) |> 
  distinct(country_name)
# Venezuela, RB does not have a label of income group.
# After double-checking The World Bank Website, this country's income level is indeed unclassified.
```

### Segmentation
```{r}
# To further focus on country and world, the cost data is segmented into two parts.
cost.country <- cost |> filter(!is.na(region)) # 217 countries
cost.region  <- cost |> filter( is.na(region)) #  48  regions / world
```

# Exploration

## Overview

```{r}
cost.region |> 
    filter(country_name == "World" & !is.na(Total_Expense)) |> 
    ggplot(aes(x = Year, y = Total_Expense, group = country_name)) +
    geom_line() +
    geom_point() +
    labs(y = "% of government overall expenditure",
         title = str_to_title("government expenditure on education"))
```

```{r echo=FALSE}
ggsave(path = here("Project 1/03. Output"), filename = "01-1. World Expenditure on Education.tiff",
       width = 8, height = 5, device = 'tiff', dpi = 500)
```

-   It seems that the expenditure on education has decreased in general.

------------------------------------------------------------------------

::: callout-caution
## Different Value

-   According to the detail provided by [the World Bank](https://reurl.cc/VRAlMQ), the value of the world was aggregated by median; however, after manually computing, the value obtained is not as same as that figure in the original data (see the following plot)

-   It is possible that the World Bank has considered something important and then added additional calculation. Or, I missed something!

-   To be compatible with other information provided by the World Bank, the value of the world provided by the World Bank would be used in the following analyses.
:::

```{r}
cost.country |> 
    filter(Year >= 2000 & Year < 2021) |>
    group_by(Year) |>
    summarise(Total_Expense = median(Total_Expense, na.rm = TRUE)) |> 
    ggplot(aes(x = Year, y = Total_Expense, group = 1)) +
    geom_line(color = "red") +
    geom_line(data = filter(cost.region, country_name == "World" & !is.na(Total_Expense)),
              aes(x = Year, y = Total_Expense, group = country_name),
              color = "black") +
    annotate("text", x = 19.5, y = 14.2, label = "my computation", size = 5, color = "red") +
    annotate("text", x = 18.7, y = 12.8, label = "given already", size = 5, color = "black") +
    labs(y = "% of government overall expenditure",
         title = str_to_title("government expenditure on education - Comparision"))
```

```{r echo=FALSE}
ggsave(path = here("Project 1/03. Output"), filename = "01-2. World Expenditure on Education (mine and world data).tiff",
       width = 8, height = 5, device = 'tiff', dpi = 500)
```

------------------------------------------------------------------------

## Q 01: Is the any difference among different income groups and regions?

### Income Groups

```{r message=FALSE}
cost.country |> 
    filter(!is.na(income_group) & Year >= 2000) |> # Exclude Venezuela
    group_by(income_group, Year) |> 
    summarise(Total_Expense = median(Total_Expense, na.rm = TRUE)) |> 
    ggplot(aes(x = Year, y = Total_Expense, 
               color = income_group, group = income_group)) +
    geom_line() +
    labs(y = "% of government overall expenditure",
         title = str_to_title("government expenditure on education by income group")) +
    geom_line(data = filter(cost.region, country_name == "World" & !is.na(Total_Expense)), 
              aes(x = Year, y = Total_Expense), color = "black") +
    annotate("text", x = 21, y = 12.4, label = "World", size = 5) +
    theme(legend.position = "bottom") +
    scale_color_discrete(name = "Income Group")
```

```{r echo=FALSE}
ggsave(path = here("Project 1/03. Output"), filename = "02. World Expenditure on Education by income group.tiff",
       width = 8, height = 5, device = 'tiff', dpi = 500)
```

-   Except for the Lower middle income group fluctuating, other 3 groups show an overall decrease in the expenditure on education. Additionally, the extent of low and upper middle income group is larger than that of high income group.

-   The "High Income" group has spend less proportion of money on education

::: callout-note
## Question(s) Raised

-   It is strange! To our impression, high income region, for example Nordic countries or Germany, emphasize people's education right; therefore, they might allocate lots of money on education system to support its operation.

-   Maybe...

    1.  The data is in percent so it is possible that the amount of money of high income spend on education is higher than that of other three income types.

    2.  These high-income countries have allocated larger proportions of government expenditure on other areas, such as social welfare.

:::

------------------------------------------------------------------------

### Region

```{r message=FALSE}
cost.country |> 
  filter(Year >= 2000) |> 
  group_by(region, Year) |> 
  summarise(Total_Expense = median(Total_Expense, na.rm = TRUE)) |> 
  ggplot(aes(x = Year, y = Total_Expense, color = region, group = region)) +
  geom_line() + 
  labs(y = "% of government overall expenditure",
       title = str_to_title("government expenditure on education by region")) +
    theme(legend.position = c(.98, .97), 
          legend.justification = c("right", "top"), 
          legend.box.just = "right", 
          legend.margin = margin(6, 6, 6, 6)) +
    scale_color_discrete(name = "Region") 
```

```{r echo=FALSE}
ggsave(path = here("Project 1/03. Output"), filename = "03. World Expenditure on Education by region.tiff",
       width = 8, height = 5, device = 'tiff', dpi = 500)
```

::: callout-note
Given that there are seven regions, which is hard to interpret for humans, they were separated into three parts based on areas.
:::

#### Africa and Middle East

```{r message=FALSE}
cost.country |> 
  filter(Year >= 2000 & !is.na(region)) |> 
  group_by(region, Year) |> 
  summarise(Total_Expense = median(Total_Expense, na.rm = TRUE)) |> 
  mutate(region_combined = case_when(region %in% c("Middle East & North Africa", "Sub-Saharan Africa") ~ "Africa",
                                     region %in% c("Latin America & Caribbean", "North America") ~ "America",
                                     TRUE ~ "Asia and Europe")) |> 
  filter(region_combined == "Africa") |> 
  ggplot(aes(x = Year, y = Total_Expense, color = region, group = region)) +
  geom_line() + 
  labs(y = "% of government overall expenditure",
       title = str_to_title("government expenditure on education in Africa")) +
    theme(legend.position = c(.98, .97), 
          legend.justification = c("right", "top"), 
          legend.box.just = "right", 
          legend.margin = margin(6, 6, 6, 6)) +
    scale_color_discrete(name = "Region")
```

```{r echo=FALSE}
ggsave(path = here("Project 1/03. Output"), filename = "04-1. World Expenditure on Education in Africa and Middle East.tiff",
       width = 8, height = 5, device = 'tiff', dpi = 500)
```

-   Compared to the proportion of government in Sub-Saharan Africa fluctuating between 14% and 18%, the figure of Middle East & North Africa showed a more substantial drop from 2000 to 2021.

#### America

```{r message=FALSE}
cost.country |> 
  filter(Year >= 2000 & !is.na(region)) |> 
  group_by(region, Year) |> 
  summarise(Total_Expense = median(Total_Expense, na.rm = TRUE)) |> 
  mutate(region_combined = case_when(region %in% c("Middle East & North Africa", "Sub-Saharan Africa") ~ "Africa",
                                     region %in% c("Latin America & Caribbean", "North America") ~ "America",
                                     TRUE ~ "Asia and Europe")) |> 
  filter(region_combined == "America") |> 
  ggplot(aes(x = Year, y = Total_Expense, color = region, group = region)) +
  geom_line() + 
  labs(y = "% of government overall expenditure",
       title = str_to_title("government expenditure on education in America")) +
    theme(legend.position = c(.98, .97), 
          legend.justification = c("right", "top"), 
          legend.box.just = "right", 
          legend.margin = margin(6, 6, 6, 6)) +
    scale_color_discrete(name = "Region")
```

```{r echo=FALSE}
ggsave(path = here("Project 1/03. Output"), filename = "04-2. World Expenditure on Education in America.tiff",
       width = 8, height = 5, device = 'tiff', dpi = 500)
```

-   In general, the expenditure on education of government in North America shifted from 11% to 14%, but the proportion of education expenditure of Latin America and Caribbean decreased.

#### Asia and Europe

```{r message=FALSE}
cost.country |> 
  filter(Year >= 2000 & !is.na(region)) |> 
  group_by(region, Year) |> 
  summarise(Total_Expense = median(Total_Expense, na.rm = TRUE)) |> 
  mutate(region_combined = case_when(region %in% c("Middle East & North Africa", "Sub-Saharan Africa") ~ "Africa",
                                     region %in% c("Latin America & Caribbean", "North America") ~ "America",
                                     TRUE ~ "Asia and Europe")) |> 
  filter(region_combined == "Asia and Europe") |> 
  ggplot(aes(x = Year, y = Total_Expense, color = region, group = region)) +
  geom_line() + 
  labs(y = "% of government overall expenditure",
       title = str_to_title("government expenditure on education in Asia and Europe")) +
    theme(legend.position = c(.98, .97), 
          legend.justification = c("right", "top"), 
          legend.box.just = "right", 
          legend.margin = margin(6, 6, 6, 6)) +
    scale_color_discrete(name = "Region") 
```

```{r echo=FALSE}
ggsave(path = here("Project 1/03. Output"), filename = "04-3. World Expenditure on Education in Asia and Europe.tiff",
       width = 8, height = 5, device = 'tiff', dpi = 500)
```

-   Governments in South Asia gradually allocated less proportion of expenditure on education between 2000 and 2021; however, from 2000 to 2020, two decades, the percent of government expenditure in East Asia & Pacific region and Europe & Central Asia on education remained static. In addition, the changing level of the former is larger than that of the latter.

------------------------------------------------------------------------

### Explore the relationship between regions and income groups

```{r}
cost.country |> 
    select(country_name, region, income_group) |> 
    distinct() |> 
    group_by(region, income_group) |> 
    count() |> 
    filter(!is.na(income_group)) |> 
    ggplot(aes(x = region, y = n, fill = income_group)) +
    geom_col(position = "fill") +
    scale_y_continuous(breaks = seq(0, 1, 0.10)) +
    scale_fill_discrete(name = "Income Group") +
    labs(x = "Region", y = "Proportion", title = str_to_title("The proportion of four income group in each region")) +
    theme(legend.position = "bottom")
```

```{r}
cost.country |> 
    select(country_name, region, income_group) |> 
    distinct() |> 
    filter(!is.na(income_group)) |> 
    tabyl(region, income_group) |> 
    as_tibble() |> rowwise() |> mutate(count = sum(c_across(`Low income`:`High income`))) |> 
    mutate(across(`Low income`:`High income`, ~ . / count)) |> 
    arrange(desc(`High income`)) |> 
    gt() |> 
    tab_header(title = "Region and Income Group (Proportion)") |> 
    fmt_number(columns = 2:5, decimals = 2) |> 
    cols_align_decimal()
```

-   In sub-Saharan Africa, half of countries are categorized as low income countries. In contrast, in Europe and central Asia, above 65% of countries is classified as high income countries.

```{r message=FALSE, warning=FALSE}
cost.country |> 
  filter(Year >= 2000 & !is.na(region) & !is.na(income_group)) |> 
  group_by(region, income_group, Year) |> 
  summarise(Total_Expense = median(Total_Expense, na.rm = TRUE)) |> 
  mutate(region_combined = case_when(region %in% c("Middle East & North Africa", "Sub-Saharan Africa") ~ "Africa",
                                     region %in% c("Latin America & Caribbean", "North America") ~ "America",
                                     TRUE ~ "Asia and Europe")) |> 
  ggplot(aes(x = Year, y = Total_Expense, color = region, group = region)) +
  geom_line() + 
  labs(y = "% of government overall expenditure",
       title = str_to_title("government expenditure on education by region and income group")) +
    theme(legend.position = "none") +
  scale_x_discrete(breaks = seq(2000, 2021, 7)) +
  facet_grid(income_group ~ region, space = "free") 
```

```{r echo=FALSE}
ggsave(path = here("Project 1/03. Output"), filename = "05. World Expenditure on Education by income and region.tiff",
       width = 8, height = 5, device = 'tiff', dpi = 500)
```




















